Reformulating phylogenetic mixed models to improve flexibility and speed

### 2018-09-05 main.tex
temp pushing

\subsection{Reformulating phylogenetic correlation matrix}
% The standard problem of phylogenetic comparative methods is to analyze relationships among data where the observations are gathered from nodes (usually tips) of a phylogenetic tree.
% Phylogenetic independent contrasts is a generalization of the paired comparisons method where contrasts are taken for each bifurcation (nodes) in a phylogenetic tree. 
% Assuming that traits evolve independently in each lineage following speciation, then the trait divergences that occur at one node are independent of divergence at other nodes.

@@

\subsubsection*{Constructing the random effects model matrix}

$\bS$ is a matrix product of an $n \times b$ indicator matrix $\bS_{id}$ of branch indices, and a branch length vector $\bB$.
For example, \mli{ref fig:tree} is a phylogenetic tree with three species:

\begin{align*}
\bS_{id} = \begin{bmatrix}
1 & 1 & 0 & 0 \\ 
1 & 0 & 1 & 0 \\ 
0 & 0 & 0 & 1
\end{bmatrix} , 
\bB = \begin{bmatrix}
\ell_1 \\
\ell_2 \\
\ell_3 \\
\ell_4 
\end{bmatrix}
\end{align*}

@@

\begin{align}
\bBadj & = (\frac{\bell}{\bomega})^{1/2} \\
\bomega & = \textrm{det}(\bSigma_{phy})^{1/n},
\end{align}
% \bmb{would like to avoid multi-letter notation if possible; multi-letter subscripts should be text, e.g. ${\mathbf B}_{\mbox{\tiny adj}}$ (using macros to define stuff in one place \ldots) Also consider $\ell$ as a replacement for $l$}
where $\bell$ is the branch length vector, $\bomega$ is a scaling factor, and $\bSigma_{phy}$ is the variance-covariance matrix of the phylogenetic tree.
For example, 
This is mathematically equivalent under BM assumption where we can reconstruct $\bC$ by matrix multiplying $\bS$ by itself.

>>>>

\subsection{Reformulating phylogenetic covariance matrix}
% The standard problem of phylogenetic comparative methods is to analyze relationships among data where the observations are gathered from nodes (usually tips) of a phylogenetic tree.
% Phylogenetic independent contrasts is a generalization of the paired comparisons method where contrasts are taken for each bifurcation (nodes) in a phylogenetic tree. 
% Assuming that traits evolve independently in each lineage following speciation, then the trait divergences that occur at one node are independent of divergence at other nodes.

@@

\subsubsection*{Constructing the species--branch random effects model matrix}

The error-free $\bS$ is a matrix product of an $n \times b$ indicator matrix $\bS_{ind}$ of branch indices, and a branch length vector $\bell$.

\begin{align*}
\bS_{ind} = \begin{bmatrix}
1 & 1 & 0 & 0 \\ 
1 & 0 & 1 & 0 \\ 
0 & 0 & 0 & 1
\end{bmatrix} , 
\bell = \begin{bmatrix}
\ell_1 \\
\ell_2 \\
\ell_3 \\
\ell_4 
\end{bmatrix}
\end{align*}
The $\bS$ matrix has a very nice property that $\bS \bS^T$ gives the variance-covariance matrix of the phylogeny.

@@

\begin{align}
\bell_{adj} & = (\frac{\bell}{\bomega})^{1/2} \\
\bomega & = \textrm{det}(\bSigma_{phy})^{1/n},
\end{align}
% \bmb{would like to avoid multi-letter notation if possible; multi-letter subscripts should be text, e.g. ${\mathbf B}_{\mbox{\tiny adj}}$ (using macros to define stuff in one place \ldots) Also consider $\ell$ as a replacement for $l$}
where $\bell$ is the branch length vector, $\bomega$ is a scaling factor, and $\bSigma_{phy}$ is the variance-covariance matrix of the phylogenetic tree.
For example, 
This is mathematically equivalent under BM assumption where we can reconstruct $\bC$ by matrix multiplying $\bS$ by itself.
